---
title: "Introduction to `dwpstat`"
author: "Evan Odell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Package Info

The UK Department for Work and Pensions (DWP) operates the [Stat-Xplore](https://stat-xplore.dwp.gov.uk/) database, containing official statistics on benefit claims, caseloads, and other relevant data. The DWP also operates a public API service, which `dwpstat` wraps. Full documentation for the DWP's API is available (here)[https://stat-xplore.dwp.gov.uk/webapi/online-help/Open-Data-API.html].

## API key and rate limits

Users need to sign up for a [Stat-Xplore account and create an API key](https://stat-xplore.dwp.gov.uk) to access the API. The API key can be set using `dwp_api_key()`, which checks for an `DWP_API_KEY` environmental variable, and allows users to set the key for a given session.

The API is rate limited, a limit that resets every hour. You can use `dwp_rate_limit()` to check current usage rates.

```{r rate-limit}
library(tibble)
dwp_rate_limit()
```


## Identify Available Data

Use `dwp_schema()` to identify available datasets, and the variables within those datasets. `dwp_schema()` accepts ID variables at any level and returns all metadata at the level below. If `dwp_schema()` is empty, returns all available data folders at the root level of the API.

Get a `tibble` of all available folders in the API:

```{r schema-a}
library(tibble)

a <- dwp_schema()

glimpse(a)
```

Get a `tibble` of metadata on all databases in the ESA folder

```{r schema-b}
b <- dwp_schema("str:folder:fesa")

glimpse(b)
```

Get a `tibble` of all variables on all databases in the ESA folder:

```{r schema-c}
c <- dwp_schema("str:database:ESA_Caseload")

glimpse(c)
```


Given their ID, you can use `dwp_scheme` to return the names of levels in group and field variables

```{r schema-d}
d <- dwp_schema("str:field:ESA_Caseload:V_F_ESA:CTDURTN")

glimpse(d)
```

Returns a `tibble` of levels for the duration options for ESA caseloads

```{r schema-e}
e <- dwp_schema("str:valueset:ESA_Caseload:V_F_ESA:CTDURTN:C_ESA_DURATION")

glimpse(e)
```

## Get data

Due to the structure of the API, actual data is returned in a list with six levels. The process for converting that data to a data frame or tibble varies considerably depending on the `column`, `row` and `wafer` fields used, and so no generic function for conversion is provided at this time.

The data structure can be difficult to work with in R; it is often unclear how to match together data labels, in the `fields` list level, with the actual data, in the `cubes` level. The example below shows a relatively simple process of assigning names to a query.

```{r}
library(purrr)
pip <- dwp_get_data(database = "str:database:PIP_Monthly",
                  measures = "str:count:PIP_Monthly:V_F_PIP_MONTHLY",
                  column = "str:field:PIP_Monthly:V_F_PIP_MONTHLY:SEX",
                  row = "str:field:PIP_Monthly:F_PIP_DATE:DATE2")

pip2 <- as.data.frame(map(pip$cubes, "values"))

pip_names <- pip$fields$items

names(pip2) <- pip_names[[2]]$labels

pip2$date <- pip_names[[1]]$labels

```

